---
  output: github_document
  always_allow_html: true
---

  <!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R packages for AMNOG analyses

Ramnog is a collection of packages for setting up pipelines for AMNOG-style HTA analyses

The aim of ramnog is that statistician/programmer can set up an AMNOG-type analyses with minimal familiarity with R.

# Packages
 
```{r, echo=FALSE}
library(kableExtra)
library(data.table)
dt <- data.table::data.table(
  Package = c("chef", "chefStats", "chefCriteria", "ramnog"),
  Description = c(
    "Core package that aids in creating endpoint specifications, builds the analysis pipeline, and coordinates the execution of the ananlyses",
    "Collection of statistical methods that are used to make summarizations or inferences for each endpoint/outcome",
    "Collection of criteria that are used in AMNOG dossiers to determin when specific analyses/statistics should be calculated",
    "Wrapper package tying ecosystem together"
  )
)

dt |>
  kable(format = "html", table.attr = "class='table table-bordered'") |>
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive")) |>
  row_spec(0, background = "lightblue")
```

# Aim

The aim of the ramnog framework is that a programmer has to write minimal code, and no programming to in order to set-up a new AMNOG-type analyses. For each study, the programmer will need to make, adjust, or check the following four types of code:

1.  The definition of each endpoint (or group of endpoints).
2.  A set of adam functions that makes any modifications to existing ADaM datasets (e.g., new age grouping in ADSL), or makes new ADaM datasets if none exist for the required output.
3.  (If needed) Define a set of criteria for when an endpoint should be included in the results. A library of these criteria are stored in the companion package {chefCriteria}
4.  A specification of the statistical functions used to summarize/analyze the data (usually found in the [chefStats](https://hta-pharma.github.io/chefStats/) package).

A core principal of the frameworks design is __modularity__. The core functionality of the framework as specefied in the code in [chef](https://hta-pharma.github.io/chef/) should change slowly, while functionality that is subject to more frequent changes are sectioned off in other packages ([chefStats](https://hta-pharma.github.io/chefStats/) and [cheCriteria](https://hta-pharma.github.io/chefCriteria/))

# Contributing
 
 We welcome contributions to the code base. Please see the [contributing vignette]("https://hta-pharma.github.io/chefCriteria/dev_git") for more information on how to contribute.

# Installation

The packages are available to install from GitHub:

```{r, eval=FALSE}
remotes::install_github("hta-pharma/ramnog")
```
