---
title: "Git Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Git Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



```{r}
library(data.table)
library(kableExtra)
```

# Ramnog ecosystem

The ramnog framework is an ecosystem of three R packages
```{r, echo=FALSE}
dt <-
  data.table::data.table(
    Package = c( "chef", "chefStats", "chefCriteria","ramnog"),
    Description = c(
      "Core package that aids in creating endpoint specifications and coordinates the execution of the package",
      "Collection statistical methods that are used to make summarizations or inferences for each endpoint/outcome",
      "Collection of criteria that are use in AMNOG dossiers to determin when specific analyses should be executed",
      "Wrapper package tying ecosystem together"
    )
  )
dt |> 
  kable(format = "html", table.attr = "class='table table-bordered'") |> 
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive")) |> 
  row_spec(0, background = "lightblue")

```


# Workflow

All packages use [Gitflow](https://www.atlassian.com/git/tutorials/comparing-workflows/gitflow-workflow) way of working. Please see the linked to resource for details on how to work in this style.

Specific for the ramnog ecosystem, each package always has the following three branches:

* `dev` For development purposes. `dev` is where completed `feature` branches should be pulled to and it should always be functional. Any uncompleted features should remain in a `feature` branch. 
* `test` testing environment. Once a new version or patch is ready, `dev` shouled be pulled into `test`, then `test` should be tested in a production environment.
* `main` Latest stable release and is what users will interact with. Pulls to `main` should only come from `test`
