---
title: "Strata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Strata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

The `stratify_by` endpoint parameter specifies the variables in the analysis data set on which the population will be sliced in the endpoint for more granular comparison on treatment arm effects. 

These are the key properties of the `stratify_by` parameter:

* `stratify_by` is a `list()` object
* The list may be empty if no stratification is needed. 
* Regardless of whether any srtata are specified (and meet the strata criteria c.f. [Criteria methods](methods_criteria.html)) or not, the endpoint statistics will always be derived for the *total* analysis population in the endpoint. The *total* can be considered a special stratum with only one level. 
* The stratum values are not required in `stratify_by`. All relevant statistics will be applied to each stratum.
* The strata variables are applied independently, i.e. strata will not be combined. So if e.g. you stratify by age group and gender, you will not have cross combination such as `SEX="F" & AGEGR="<65"`. 

If cross combinations of two stratum variables `A` and `B` are required, either: 

1. Create a derived variable in `data_prepare` that combines `A` and `B` and stratify on this new variable.
2. Use `endpoint_filter` or `group_by` to group or filter the data on `A` and then stratify on `B` within each group level.

# Examples

##### Ex 5.1


```{r, eval = FALSE}
# Example of partial endpoint specification with a strata definition
ep_spec_ex5_2  <- chef::mk_endpoint_str(
  stratify_by = list(c("SEX")),
  ...)
```

##### Ex 5.2

Two strata set:

```{r, eval = FALSE}
# Example of partial endpoint specification with a strata definition
ep_spec_ex5_2  <- chef::mk_endpoint_str(
  stratify_by = list(c("SEX", "AGEGGR1")),
  ...)
```

##### Ex 5.3

No strata, in which case only the *total* will be present in the endpoint:

```{r, eval = FALSE}
# Example of partial endpoint specification with a strata definition
ep_spec_ex5_3  <- chef::mk_endpoint_str(
  stratify_by = list(),
  ...)
```
