---
title: "Endpoint Label"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Endpoint Label}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

For each endpoint the user has the option to specify a label that describes the content of the endpoint. 
This is parsed as a string via the `endpoint_label` parameter in the endpoint specification.
The string may contain references to other parameters in the endpoint specification encapsulated by `<>`. 
These references will be dynamically evaluated to the respective values of the parameters.
The endpoint label does not have to be unique for each endpoint, and references are optional.

Label referencing is particularly useful when endpoint grouping is applied, which resolves in an endpoint for each group level. 
In this case you may dynamically refer to the group level present in each of these endpoints.

The dynamic labeling may also be a convenient way to refer to other important parameters that identifies each endpoint e.g., `pop_var`, `period_var` and `endpoint_filter`.


# Examples


##### Ex 6.1

Suppose you define an endpoint specification with an endpoint on the safety analysis population (`SAFFL = "Y"`) grouped on adverse event severity `AESEV` which has three levels: `MILD`, `MODERATE`, and `SERVERE`. 
Note that this assumes that `SAFFL` and `AESEV` are variables in the analysis data set. For this setup you may specify a dynamic endpoint label that refer to both the population and the severity: 

```{r, eval=FALSE}
# Example of partial endpoint specification of analysis population
ep_spec_ex6_1  <- chef::mk_endpoint_str(pop_var = "SAFFL",
                                 pop_value = "Y",
                                 group_by = list(list(AESEV = c())),
                                 endpoint_label = "Example: <pop_var> - <AESEV> adverse events",
                                 ...)
```

This will resolve in three endpoints with the following labels:

```{r, eval=FALSE}
"Example: SAFFL - MILD adverse events"
"Example: SAFFL - MODERATE adverse events"
"Example: SAFFL - SERVERE adverse events"
```

##### Ex 6.2

In extension of the previous example, suppose you only want to consider subjects of ages between 18-64 years via the age group variable `AGEGR1` in the analysis data set.
In addition, you want to update the endpoint label to refer to this filter. This could look like this:   

```{r, eval=FALSE}
# Example of partial endpoint specification of analysis population
ep_spec_ex6_2  <- chef::mk_endpoint_str(
  pop_var = "SAFFL",
  pop_value = "Y",
  endpoint_filter = "AGEGR1 == '18-64'",
  group_by = list(list(AESEV = c())),
  endpoint_label =
    "Example: <pop_var> - <AESEV> adverse events / <endpoint_filter>",
  ...
)
```

Again, this resolve in three endpoints and they will now have these endpoint labels:

```{r, eval=FALSE}
"Example: SAFFL - MILD adverse events / AGEGR1 == '18-64'"
"Example: SAFFL - MODERATE adverse events / AGEGR1 == '18-64'"
"Example: SAFFL - SERVERE adverse events / AGEGR1 == '18-64'"
```

##### Ex 6.3

If you do not provide any references in the endpoint label, it is simply returned as a static string.
Hence, if you modify the label in the previous example as follows:

```{r, eval=FALSE}
# Example of partial endpoint specification of analysis population
ep_spec_ex6_3  <- chef::mk_endpoint_str(
  pop_var = "SAFFL",
  pop_value = "Y",
  endpoint_filter = "AGEGR1 == '18-64'",
  group_by = list(list(AESEV = c())),
  endpoint_label = "Example of a fixed endpoint label",
  ...
)
```

You will get the following endpoint labels:

```{r, eval=FALSE}
"Example of a fixed endpoint label"
"Example of a fixed endpoint label"
"Example of a fixed endpoint label"
```
