---
title: "Quality control for ramnog ecosystem"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quality control for ramnog ecosystem}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(ramnog)
library(testthat)
library(data.table)
library(chef)
library(testr)
library(chefStats)
```

The validity of the ramnog framework is ensured on several levels:

 - Unit test individual functions/modules in all packages
 - Integration/end-to-end testing to ensure the interactions between individual functions/modules (that have been unit tested) are as expected
 - CI/CD to ensure changes are properly tested and deployed to the users

## Unit testing

Each package contains the unit tests for the functionality that package is responsible for. This means, for example, it is the responsibility of the unit-tests in {chefStats} to ensure that the statistical functions written in chefStats produce the result they are supposed to produce.


## Integration testing

Integration testing verifies the framework produces the expected results

The integration testing is done via the {testthat} framework, and are essentially unit-tests that cover many different units

The goal is to compare the results of an analysis run via ramnog with the expected results done from a manual calculation of the same analysis.

For example, below is the unit test to ensure the results of an endpoint definition defined and run via chef match those calculated manually

<details>
<summary>Example of integration test using {testthat} framework </summary>

```{r}
testthat::test_that("Valide stats when one strata level is not found",
          {
            # SETUP -------------------------------------------------------------------
            testr::create_local_project()
            mk_adae <- function(study_metadata) {
              adsl <- pharmaverseadam::adsl |> data.table::setDT()
              adae <- pharmaverseadam::adae |> data.table::setDT()
              adae_out <-
                merge(adsl, adae[, c(setdiff(names(adae), names(adsl)), "USUBJID"), with =
                                   F], by = "USUBJID", all = TRUE)
              adae_out[TRT01A %in% c('Placebo', 'Xanomeline High Dose'),]
            }
            mk_ep_def <- function() {
              chef::mk_endpoint_str(
                study_metadata = list(),
                pop_var = "SAFFL",
                pop_value = "Y",
                treatment_var = "TRT01A",
                treatment_refval = "Xanomeline High Dose",
                stratify_by = list(c("SEX")),
                data_prepare = mk_adae,
                endpoint_label = "A",
                custom_pop_filter = "SEX == 'F'",
                group_by = list(list(AESEV = c())),
                stat_by_strata_by_trt = list(
                  "N" = chefStats::n_subj,
                  "E" = chefStats::n_event,
                  "N subj E" = chefStats::n_subj_event
                ),
                stat_by_strata_across_trt = list("RR" = chefStats::RR,
                                                 "OR" = chefStats::OR),
                stat_across_strata_across_trt = list("P-interaction" = chefStats::p_val_interaction)
              )
            }

            chef::use_chef(
              pipeline_dir = "pipeline",
              r_functions_dir = "R/",
              pipeline_id = "01",
              mk_endpoint_def_fn = mk_ep_def,
              mk_adam_fn = list(mk_adae)
            )
            # ACT ---------------------------------------------------------------------
            targets::tar_make()
            targets::tar_load(ep_stat)
            # EXPECT ------------------------------------------------------------------
<<<<<<< HEAD
            # Manually calculated expected results
            x <- mk_adae(study_metadata = NULL)
=======
            x <- mk_adae_demo()
>>>>>>> 62e189b (Adds junit as reporter)
            # Relative Risk
            actual <-
              ep_stat[endpoint_group_filter == "AESEV == \"SEVERE\"" &
                        fn_name == "RR" &
                        stat_filter == "SEX == \"F\"" & stat_result_label == "RR", stat_result_value]
            x1 <- x[SEX == "F" & SAFFL == "Y"]
            x1[, event := FALSE]
            x1[AESEV == "SEVERE", event := TRUE] |> data.table::setorder(-event)
            x1_unique <-
              unique(x1, by = "USUBJID", fromLast = FALSE)
            two_by_two <-
              x1_unique[, .N, by = .(TRT01A, event)][order(TRT01A, event)]
            
            # Outcome, exposed
            a <-
              two_by_two[TRT01A == "Xanomeline High Dose" &
                           (event), N] 
            
            # No outcome, exposed
            b <-
              two_by_two[TRT01A == "Xanomeline High Dose" &
                           !(event), N] 
             
            # Outcome, not exposed
            c <-
              two_by_two[TRT01A == "Placebo" &
                           (event), N]
            
            # No outcome, not exposed
            d <-
              two_by_two[TRT01A == "Placebo" &
                           !(event), N] 
            
            expected <- (a / sum(a, b)) / (c / sum(c, d))
            expect_identical(actual, expected)

            # Number of Events
            actual <-
              ep_stat[endpoint_group_filter == "AESEV == \"SEVERE\"" &
                        fn_name == "E" &
                        strata_var == "SEX"]
            expected <-
              x1[(event), .N, by = .(TRT01A)][order(TRT01A)][, as.double(N)]
            expect_identical(actual$stat_result_value, expected)
          })
```
</details>

## CI/CD 
To be added
